\section*{Data availability}

All source datasets used in this study are publicly available. TCGA PanCancer Atlas and TCGA-BRCA molecular and clinical data were accessed through cBioPortal. METABRIC, CPTAC 2020, and SMC 2018 molecular and clinical data were accessed through cBioPortal where available. GSE96058/SCAN-B data were accessed through the Gene Expression Omnibus and the associated public processed data package. Public datasets and metadata were accessed and processed through 14 July 2026. A peer-review reproducibility archive containing analysis scripts, processed result tables, generated figures, and environment information is provided as supplementary material. A public GitHub repository and Zenodo archive DOI will be added before final publication.

\section*{Code availability}

The supplementary reproducibility archive contains processed result tables required to reproduce manuscript tables and figures, along with Python and package-version information. The main scripts in the archive include:
\begingroup
\footnotesize
\begin{itemize}
\item \path{scripts/analyze_alignment_domain_shift_benchmark.py}
\item \path{scripts/reviewer_cv_noise_significance_analysis.py}
\item \path{scripts/strengthen_alignment_liquid_medical_evidence.py}
\item \path{scripts/analyze_multicancer_reliability_extension.py}
\item \path{scripts/expand_multicancer_clinical_scope.py}
\end{itemize}
\endgroup

\section*{Author contributions}

Z.H. conceived the study, designed the computational experiments, processed the data, implemented the analysis pipeline, performed model training and evaluation, prepared figures and tables, and drafted the manuscript. X.B. supervised the study, contributed clinical interpretation, and revised the manuscript. Both authors reviewed and approved the final manuscript.

\section*{Competing interests}

The authors declare that they have no competing interests.

\section*{Funding}

No specific funding was received for this work.

\section*{Use of generative artificial intelligence}

Generative artificial intelligence tools were used to assist with manuscript drafting, language editing, code organization, and internal consistency checks. All analyses, interpretations, and manuscript content were reviewed and approved by the authors, who take full responsibility for the final work. No generative artificial intelligence tool is listed as an author.
